---
title: "RS07.sPLS_1mon"
author: "Jay"
date: "2023-11-28"
output: html_document
---

In this part, sPLS model on gut microbiome, environmental factors and both of them predicting 1-year CS score as continuous score will be performed

The output is expected to be (ROC curve) correlation plot in this case, to illustrate the predictive power of gut microbiome and environmental factor individually and jointly.
The aim of this part is to investigate how strong the early gut mcirobime and envrionmental factors will impact the gut microbiome status at 1 year timepoint.


```{r setup, include=FALSE}

library(caret)
library(pROC)
library(mixOmics)
library(tidyverse)
library(mixOmicsCaret) 
library(ggforce)
library(magrittr)
library(copiome)
library(dplyr)
library(phyloseq)
```


###set otu res path
```{r}
outpath <- "../results/RS07_01.sPLS_1mon"
dir.create(outpath)
```

# Env VS CS score in full dataset
## Load data.
```{r}

load("../data/phy_3timepoints_restoration_score.rdata")

phy_1mon <- phy_3timepoints %>% subset_samples(Visit == "1 month" & !restoration_score %in% c(NA))
samp <- sample_df(phy_1mon)
colid = which(colnames(samp) == "ABCNO")
col1 = which(colnames(samp) == "delivery")
col2 = which(colnames(samp) == "restoration_score")

samp_pls <- sample_df(phy_1mon) %>% .[,c(colid,col1:col2)] %>% dplyr::select(-sibling_age,-sibling_age_imputed,-CS_Preds_PLS_1y) %>% na.omit()


```



```{r}
set.seed(123)
library(mixOmicsCaret)
library(tidyr)
library(caret)

repCV10 <- trainControl(method = "repeatedcv", 
                        number = 10, 
                        repeats = 11, 
                        returnResamp = "all", 
                        savePredictions = "all", 
                        allowParallel = T, 
                        verboseIter = F)

```


###sPLS 1
```{r}
set.seed(123)
scorePLS <- train(restoration_score ~ ., data = samp_pls %>% dplyr::select(-ABCNO),
                  method = get_mixOmics_spls(),
                  preProc = c("center", "scale"),
                  tuneGrid = expand.grid(ncomp = 1:5, 
                                         keepX = c(1,2,4,6,8,10,22,25,30,35,39), 
                                         keepY = 1),
                  trControl = repCV10,
                  fixX = c())

saveRDS(scorePLS, file = paste0(outpath, "/sPLS_1mon_env.rds"))

```

```{r}
scorePLS <- readRDS("../results/RS07_01.sPLS_1mon/sPLS_1mon_env.rds")
p <- scorePLS$pred %>%
  separate(Resample, c("Fold", "Rep")) %>% filter(ncomp == 1) %>%  
  dplyr::group_by(ncomp,keepX, Rep) %>%
  dplyr::summarise(cor = cor(obs, pred, method = "spearman")) %>%
  ggplot(aes(x = factor(keepX), y = cor)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(color = "grey", alpha = 0.8, show.legend = FALSE) +
    facet_grid(. ~ ncomp, labeller = as_labeller(c(`1` = "Component 1"))) +
    geom_hline(yintercept = 0) +
    ylim(c(-0.15,0.3))+
    scale_color_brewer(palette = "Set1", name = NULL) +
    theme_bw() + theme(strip.background = element_blank()) +
    xlab("Number of variables")+
    ylab("Spearman coefficient of true value VS predictions");p

plotname1 = paste(outpath,"/sPLS_CV_corplot_Env_1mon.pdf",sep = "")
ggsave(plotname1, p,width = 8,height = 5,limitsize = FALSE)


save(p, file = paste0(outpath, "/sPLS_CV_corplot_Env_1mon.RData"))
```



##get predictions
###Retrieve the predictions from the optimal model
```{r}
# Retrieve the predictions from the optimal model
bestSet <- scorePLS %>% get_best_predictions %>% group_by(Rep) 
head(bestSet)

bestSet %>% 
  summarize(cor = cor(obs, pred, method = "spearman") %>% as.numeric) %>% 
  arrange(cor)

savedPreds <- scorePLS %>% get_best_predictions %>% filter(Rep == "Rep05")
head(savedPreds)


cor.cv <- cor.test(samp_pls$restoration_score,savedPreds$pred,method = "spearman", use = "complete.obs")
cor <- cor.cv$estimate %>% round(3)

df <- scorePLS %>% get_loadings("finalModel") %>% 
  filter(loading != 0)


# Creating the group_info data frame
group_info <- data.frame(
  var = c(
    "atb_birth_motheryes",
    "oldchild_ynyes"
  ),
  readable_label = c(
    "Antibiotics to mother at birth (yes)",
    "Having older siblings (yes)"
  )
)

df <- df %>% merge(group_info, by = "var")


p <-  df %>% 
  ggplot(aes(reorder(readable_label, loading), loading)) +
  geom_col(fill = "grey",show.legend = FALSE) +
  ylim(c(-1,1))+
  coord_flip()+
  facet_grid(. ~ comp, labeller = as_labeller(c(`comp1` = "Environmental factors"))) +

  xlab(NULL)+
  ylab("Loading")+
  theme_minimal()+
  theme(axis.text.x = element_text(size = 14,angle = 75, hjust = 1),
        axis.text.y = element_text(size = 14),
        strip.background = element_rect(fill = "grey", color = "grey"),
        strip.text.x = element_text(size = 14),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank());p

plotname1 = paste(outpath,"/loading_Env_vs_restoration_score.pdf",sep = "")
ggsave(plotname1, p,width = 5,height = 3,limitsize = FALSE)

save(p, file = paste0(outpath, "/loading_Env_vs_restoration_score.RData"))


# CSscore_Preds_env <- full_scores$predict %>% as.data.frame()
RScore_Preds_env_1mon <- savedPreds$pred %>% as.data.frame()

colnames(RScore_Preds_env_1mon) <- c("RScore_Preds_env_1mon")
RScore_Preds_env_1mon$ABCNO <- samp_pls$ABCNO

tablename <- paste(outpath,"/RScore_Preds_env_1mon",".csv",sep = "")
write.csv(RScore_Preds_env_1mon,tablename)

```








# Gut microbiome VS CS score in full dataset
##load data
```{r}

load("../data/phy_3timepoints_restoration_score.rdata")

ps <- phy_3timepoints %>% subset_samples(Visit == "1 month" & !restoration_score %in% c(NA)) %>% speedyseq::tax_glom("species") %>% filter_phy(prevalence = .1, abundance = 1e-4,compositional = FALSE) %>% transform_sample_counts(function(x) log10(x + min(x[x > 0])))

name_clean <- tax_df(ps)$species  %>% gsub("\\[|\\]", "", .) %>% gsub(" ","_",.) %>% gsub("-","_",.) %>% gsub("^_","",.) %>% gsub("^[0-9]","a\\",.)
tax_table(ps) <- as(tax_table(ps), "matrix") %>% as.data.frame() %>% mutate(species = name_clean) %>% as(., "matrix")

otu <- otu_df(ps)
colnames(otu) <- tax_df(ps)$species[match(colnames(otu),row.names(tax_df(ps)))]

feature_name <- colnames(otu)

#calibrate for sequence depth
otu$SampleID <- row.names(otu)

samp_otu <- left_join(sample_df(ps), otu)

otu.residual <- list()

for (i in feature_name){
  lm <- lm(samp_otu[,which(colnames(samp_otu) == i)] ~ samp_otu$Reads + log(samp_otu$Reads))
  otu.residual[[i]] <- residuals(lm) %>% scale
}

otu.residual <- as.data.frame(otu.residual)

#merge with sam data
otu.residual$SampleID <- samp_otu$SampleID

samp_pls_bac <- left_join(sample_df(ps), otu.residual) %>% dplyr::select(ABCNO,restoration_score, all_of(feature_name))

```

```{r}
set.seed(123)
library(mixOmicsCaret)
library(tidyr)
library(caret)

repCV10 <- trainControl(method = "repeatedcv", 
                        number = 10, 
                        repeats = 11, 
                        returnResamp = "all", 
                        savePredictions = "all", 
                        allowParallel = T, 
                        verboseIter = F)

```


##sPLS 1
```{r}

set.seed(123)
scorePLS <- train(restoration_score ~ ., data = samp_pls_bac %>% dplyr::select(-ABCNO),
                  method = get_mixOmics_spls(),
                  preProc = c("center", "scale"),
                  tuneGrid = expand.grid(ncomp = 1:5, 
                                         keepX = c(1,2,4,8, 25, 40, 60,100, 143,151), 
                                         keepY = 1),
                  trControl = repCV10,
                  fixX = c())

```


```{r}

p <- scorePLS$pred %>%
  separate(Resample, c("Fold", "Rep")) %>%
  filter(ncomp %in% c(1,2)) %>%
  group_by(ncomp, keepX, Rep) %>%
  summarise(cor = cor(obs, pred, method = "spearman")) %>%
  ggplot(aes(x = factor(keepX), y = cor)) +
    geom_boxplot(outlier.shape = NA) +
    ylim(c(-0.15,0.3))+
    facet_grid(. ~ ncomp, labeller = as_labeller(c(`1` = "Component 1", `2` = "Component 2"))) +
    geom_jitter(color = "#0072B2", alpha = 0.8, show.legend = FALSE) +
    geom_hline(yintercept = 0) +
    scale_color_brewer(palette = "Set1", name = NULL) +
    theme_bw() + 
    theme(strip.background = element_blank()) +
    xlab("Number of variables") +
    ylab("Spearman coefficient of true value VS predictions");p

plotname1 = paste(outpath,"/sPLS_CV_corplot_Bac_1mon.pdf",sep = "")
ggsave(plotname1, p,width = 8,height = 5,limitsize = FALSE)


save(p, file = paste0(outpath, "/sPLS_CV_corplot_Bac_1mon.RData"))

```

##sPLS 2
```{r}
set.seed(123)
scorePLS_2 <- train(restoration_score ~ ., data = samp_pls_bac %>% dplyr::select(-ABCNO),
                  method = get_mixOmics_spls(),
                  preProc = c("center", "scale"),
                  tuneGrid = expand.grid(ncomp = 1, 
                                         keepX = c(1,2,4,8, 25, 40, 60,100, 143,151), 
                                         keepY = 1),
                  trControl = repCV10,
                  fixX = c())
```


```{r}
p <- scorePLS_2$pred %>%
  separate(Resample, c("Fold", "Rep")) %>%
  filter(ncomp %in% c(1)) %>%
  group_by(ncomp, keepX, Rep) %>%
  summarise(cor = cor(obs, pred, method = "spearman")) %>%
  ggplot(aes(x = factor(keepX), y = cor)) +
    geom_boxplot(outlier.shape = NA) +
    ylim(c(-0.15,0.3))+
    facet_grid(. ~ ncomp, labeller = as_labeller(c(`1` = "Component 1", `2` = "Component 2"))) +
    geom_jitter(color = "#0072B2", alpha = 0.8, show.legend = FALSE) +
    geom_hline(yintercept = 0) +
    scale_color_brewer(palette = "Set1", name = NULL) +
    theme_bw() + 
    theme(strip.background = element_blank()) +
    xlab("Number of variables") +
    ylab("Spearman coefficient of true value VS predictions");p
```



##sPLS 3
```{r}
set.seed(123)
scorePLS_3 <- train(restoration_score ~ ., data = samp_pls_bac %>% dplyr::select(-ABCNO),
                  method = get_mixOmics_spls(),
                  preProc = c("center", "scale"),
                  tuneGrid = expand.grid(ncomp = 1, 
                                         keepX = c(8:40), 
                                         keepY = 1),
                  trControl = repCV10,
                  fixX = c())

saveRDS(scorePLS_3, file = paste0(outpath, "/sPLS_1mon_bac.rds"))
```


```{r}

scorePLS_3 <- readRDS("../results/RS07_01.sPLS_1mon/sPLS_1mon_bac.rds")
p <- scorePLS_3$pred %>%
  separate(Resample, c("Fold", "Rep")) %>%
  filter(ncomp %in% c(1)) %>%
  group_by(ncomp, keepX, Rep) %>%
  summarise(cor = cor(obs, pred, method = "spearman")) %>%
  ggplot(aes(x = factor(keepX), y = cor)) +
    geom_boxplot(outlier.shape = NA) +
    ylim(c(-0.15,0.3))+
    facet_grid(. ~ ncomp, labeller = as_labeller(c(`1` = "Component 1"))) +
    geom_jitter(color = "#0072B2", alpha = 0.8, show.legend = FALSE) +
    geom_hline(yintercept = 0) +
    scale_color_brewer(palette = "Set1", name = NULL) +
    theme_bw() + 
    theme(strip.background = element_blank()) +
    xlab("Number of variables") +
    ylab("Spearman coefficient of true value VS predictions");p
```


##get predictions
###Retrieve the predictions from the optimal model
```{r}
# Retrieve the predictions from the optimal model
bestSet <- scorePLS_3 %>% get_best_predictions %>% group_by(Rep) 
head(bestSet)
bestSet %>% 
  summarize(cor = cor(obs, pred, method = "spearman") %>% as.numeric) %>% 
  arrange(cor)


savedPreds <- scorePLS_3 %>% get_best_predictions %>% filter(Rep == "Rep02")
head(savedPreds)


cor.cv <- cor.test(samp_pls_bac$restoration_score,savedPreds$pred,method = "spearman",use = "complete.obs")
cor <- cor.cv$estimate %>% round(3)
df <- scorePLS_3 %>% get_loadings("finalModel") %>% 
  filter(loading != 0) %>% 
  mutate(var = gsub("_", " ", var))





p <-  df %>% 
  ggplot(aes(reorder(var,loading), loading)) +
  geom_col(fill = "#0072B2",show.legend = FALSE) +
  ylim(c(-1,1))+
  coord_flip()+
  facet_wrap(~ comp, scales = "free_x",nrow = 1, labeller = as_labeller(c(`comp1` = "1 month gut microbiome"))) +
  xlab(NULL)+
  ylab("Loading")+
  theme_minimal()+
  theme(axis.text.x = element_text(size = 14,angle = 75, hjust = 1),
        axis.text.y = element_text(size = 14,face = "italic"),
        strip.background = element_rect(fill = "#0072B280", color = "#0072B280"),
        strip.text.x = element_text(size = 14),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) ;p

plotname1 = paste(outpath,"/loading_Bac_vs_resotration_score_comp1.pdf",sep = "")
ggsave(plotname1, p,width = 8,height = 5,limitsize = FALSE)
save(p, file = paste0(outpath, "/loading_Bac_vs_restoration_score_comp1.RData"))


RScore_Preds_bac_1mon <- savedPreds$pred %>% as.data.frame()

colnames(RScore_Preds_bac_1mon) <- c("RScore_Preds_bac_1mon")
RScore_Preds_bac_1mon$ABCNO <- samp_pls_bac$ABCNO

tablename <- paste(outpath,"/RScore_Preds_bac_1mon_comp1",".csv",sep = "")
write.csv(RScore_Preds_bac_1mon,tablename)

```











# Gut microbiome plus Env VS CS score in full dataset
##load data
```{r}

load("../data/phy_3timepoints_restoration_score.rdata")

ps <- phy_3timepoints %>% subset_samples(Visit == "1 month" & !restoration_score %in% c(NA)) %>% speedyseq::tax_glom("species") %>% filter_phy(prevalence = .1, abundance = 1e-4,compositional = FALSE) %>% transform_sample_counts(function(x) log10(x + min(x[x > 0])))

name_clean <- tax_df(ps)$species  %>% gsub("\\[|\\]", "", .) %>% gsub(" ","_",.) %>% gsub("-","_",.) %>% gsub("^_","",.) %>% gsub("^[0-9]","a\\",.)
tax_table(ps) <- as(tax_table(ps), "matrix") %>% as.data.frame() %>% mutate(species = name_clean) %>% as(., "matrix")

otu <- otu_df(ps)
colnames(otu) <- tax_df(ps)$species[match(colnames(otu),row.names(tax_df(ps)))]

feature_name <- colnames(otu)

#calibrate for sequence depth
otu$SampleID <- row.names(otu)

samp_otu <- left_join(sample_df(ps), otu)

otu.residual <- list()

for (i in feature_name){
  lm <- lm(samp_otu[,which(colnames(samp_otu) == i)] ~ samp_otu$Reads + log(samp_otu$Reads))
  otu.residual[[i]] <- residuals(lm) %>% scale
}

otu.residual <- as.data.frame(otu.residual)

#merge with sam data
otu.residual$SampleID <- samp_otu$SampleID

samp_name <- sample_df(ps) %>% .[,c(col1:col2)] %>% dplyr::select(-sibling_age,-sibling_age_imputed,-CS_Preds_PLS_1y) %>% colnames


samp_pls_env_bac <- left_join(sample_df(ps), otu.residual) %>% dplyr::select(ABCNO,all_of(samp_name), all_of(feature_name)) %>% na.omit()

```

```{r}
set.seed(123)
library(mixOmicsCaret)
library(tidyr)
library(caret)
repCV10 <- trainControl(method = "repeatedcv", 
                        number = 10, 
                        repeats = 11, 
                        returnResamp = "all", 
                        savePredictions = "all", 
                        allowParallel = T, 
                        verboseIter = F)

```


##sPLS 1
```{r}

set.seed(123)
scorePLS <- train(restoration_score ~ ., data = samp_pls_env_bac %>% dplyr::select(-ABCNO),
                  method = get_mixOmics_spls(),
                  preProc = c("center", "scale"),
                  tuneGrid = expand.grid(ncomp = 1:5, 
                                         keepX = c(1,2,4,6,8,10,20,30,50,70,80,120), 
                                         keepY = 1),
                  trControl = repCV10,
                  fixX = c())

```


```{r}
scorePLS$pred %>%
  separate(Resample, c("Fold", "Rep")) %>%
  dplyr::group_by(ncomp, keepX, Rep) %>%
  dplyr::summarise(cor = cor(obs, pred, method = "spearman")) %>%
  ggplot(aes(x = factor(keepX), y = cor)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(aes(color = factor(ncomp)), alpha = 0.2, show.legend = FALSE) +
    geom_hline(yintercept = 0) +
    facet_grid(. ~ ncomp) +
    scale_color_brewer(palette = "Set1", name = NULL) +
    theme_bw() + theme(strip.background = element_blank()) +
    xlab("Number of variables")
```

##sPLS 2
```{r}
set.seed(123)
scorePLS_2 <- train(restoration_score ~ ., data = samp_pls_env_bac %>% dplyr::select(-ABCNO),
                  method = get_mixOmics_spls(),
                  preProc = c("center", "scale"),
                  tuneGrid = expand.grid(ncomp = 1, 
                                         keepX = c(1:40), 
                                         keepY = 1),
                  trControl = repCV10,
                  fixX = c())
```


```{r}
scorePLS_2$pred %>%
  separate(Resample, c("Fold", "Rep")) %>%
  dplyr::group_by(ncomp, keepX, Rep) %>%
  dplyr::summarise(cor = cor(obs, pred, method = "spearman")) %>%
  ggplot(aes(x = factor(keepX), y = cor)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(aes(color = factor(ncomp)), alpha = 0.2, show.legend = FALSE) +
    geom_hline(yintercept = 0) +
    facet_grid(. ~ ncomp) +
    scale_color_brewer(palette = "Set1", name = NULL) +
    theme_bw() + theme(strip.background = element_blank()) +
    xlab("Number of variables")


p <- scorePLS_2$pred %>%
  separate(Resample, c("Fold", "Rep")) %>% filter(ncomp %in% c(1,2)) %>%  
  dplyr::group_by(ncomp,keepX, Rep) %>%
  dplyr::summarise(cor = cor(obs, pred, method = "spearman")) %>%
  ggplot(aes(x = factor(keepX), y = cor)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(aes(color = "#D55E00", alpha = 0.2), show.legend = FALSE) +
    facet_grid(. ~ ncomp, labeller = as_labeller(c(`1` = "Component 1"))) +
    geom_hline(yintercept = 0) +
    scale_color_brewer(palette = "Set1", name = NULL) +
    theme_bw() + theme(strip.background = element_blank()) +
    xlab("Number of variables")+
    ylab("Spearman coefficient of true value VS predictions");p


plotname1 = paste(outpath,"/sPLS_CV_corplot_Env&Bac_1mon.pdf",sep = "")
ggsave(plotname1, p,width = 8,height = 5,limitsize = FALSE)


save(p, file = paste0(outpath, "/sPLS_CV_corplot_Env&Bac_1mon.RData"))
```



##sPLS 3
```{r}
set.seed(123)
scorePLS_3 <- train(restoration_score ~ ., data = samp_pls_env_bac %>% dplyr::select(-ABCNO),
                  method = get_mixOmics_spls(),
                  preProc = c("center", "scale"),
                  tuneGrid = expand.grid(ncomp = 1, 
                                         keepX = c(1:40), 
                                         keepY = 1),
                  trControl = repCV10,
                  fixX = c(17))

saveRDS(scorePLS_3, file = paste0(outpath, "/sPLS_1mon_bac_env.rds"))
```


```{r}
scorePLS_3 <- readRDS("../results/RS07_01.sPLS_1mon/sPLS_1mon_bac_env.rds")
p <- scorePLS_3$pred %>%
  separate(Resample, c("Fold", "Rep")) %>%
  dplyr::group_by(ncomp, keepX, Rep) %>%
  dplyr::summarise(cor = cor(obs, pred, method = "spearman")) %>%
  ggplot(aes(x = factor(keepX), y = cor)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(aes(color = factor(ncomp)), alpha = 0.2, show.legend = FALSE) +
    geom_hline(yintercept = 0) +
    facet_grid(. ~ ncomp) +
    scale_color_brewer(palette = "Set1", name = NULL) +
    theme_bw() + theme(strip.background = element_blank()) +
    xlab("Number of variables");p

plotname1 = paste(outpath,"/sPLS_CV_corplot_Env&Bac_1mon_17_comp1.pdf",sep = "")
ggsave(plotname1, p,width = 8,height = 5,limitsize = FALSE)


save(p, file = paste0(outpath, "/sPLS_CV_corplot_Env&Bac_1mon_17_comp1.RData"))

```




##get predictions
###Retrieve the predictions from the optimal model
```{r}
# Retrieve the predictions from the optimal model
bestSet <- scorePLS_3 %>% get_best_predictions %>% group_by(Rep)
head(bestSet)
bestSet %>%
  summarize(cor = cor(obs, pred, method = "spearman") %>% as.numeric) %>%
  arrange(cor)


savedPreds <- scorePLS_3 %>% get_best_predictions %>% filter(Rep == "Rep04")
head(savedPreds)


cor.cv <- cor.test(samp_pls_env_bac$restoration_score,savedPreds$pred,method = "spearman", use = "complete.obs")
cor <- cor.cv$estimate %>% round(3)


df <- scorePLS_3 %>% get_loadings("finalModel") %>% 
  mutate(type = if_else(var %in% feature_name, "bac", "env")) %>% 
  filter(loading != 0) 

env_names <- df %>% filter(type == "env") %>% .$var


group_info <- data.frame(
  var = c(
    "deliveryc_section",
    "atb_birth_motheryes",
    "oldchild_ynyes"
  ),
  readable_label = c(
    "Delivery by C-section",
    "Antibiotics to mother at birth",
    "Having older siblings (yes)"
  ))

index = match(df$var, group_info$var,nomatch = FALSE )

df[index,]$var <- group_info$readable_label

df[!index, ]$var <- paste0("*", gsub("_", " ", df[!index, ]$var), "*")

df <- df %>% arrange(loading)

library(ggtext)
p <- df %>%
  ggplot(aes(reorder(var, loading), loading, fill = type)) +
  geom_col(show.legend = FALSE) +
    coord_flip()+

  scale_fill_manual(values = c("env" = "grey", "bac" = "#0072B2")) +
  ylim(c(-1, 1)) +
  facet_wrap(~ comp, scales = "free_x", nrow = 1, labeller = as_labeller(c('comp1' = "1 month gut microbiome and environmental factors"))) +
  xlab(NULL) +
  ylab("Loading") +
  theme_minimal() +
  theme(
    axis.text.x = element_markdown(size = 14,angle = 75, hjust = 1),
    axis.text.y = element_markdown(size = 14),
    strip.background = element_rect(fill = "#D55E0080", color = "#D55E0090"),
    strip.text.x = element_text(size = 14),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank()) +
  scale_x_discrete(labels = df$var);p

plotname1 = paste(outpath,"/loading_Bac_Env_vs_restoration_score.pdf",sep = "")
ggsave(plotname1, p,width = 8,height = 5,limitsize = FALSE)


save(p, file = paste0(outpath, "/loading_Bac_Env_vs_restoration_score.RData"))


Rscore_Preds_env_bac_1mon <- savedPreds$pred %>% as.data.frame()

colnames(Rscore_Preds_env_bac_1mon) <- c("Rscore_Preds_env_bac_1mon")
Rscore_Preds_env_bac_1mon$ABCNO <- samp_pls_env_bac$ABCNO

tablename <- paste(outpath,"/Rscore_Preds_env_bac_1mon",".csv",sep = "")
write.csv(Rscore_Preds_env_bac_1mon,tablename)
```


# Summary
```{r}
Rscore_Preds_env_bac_1mon <- rio::import("../results/RS07.sPLS_1mon/Rscore_Preds_env_bac_1mon.csv") %>% dplyr::select(-V1)
Rscore_Preds_env_1mon <- rio::import("../results/RS07.sPLS_1mon/RScore_Preds_env_1mon.csv") %>% dplyr::select(-V1)
Rscore_Preds_bac_1mon <- rio::import("../results/RS07.sPLS_1mon/RScore_Preds_bac_1mon_comp1.csv") %>% dplyr::select(-V1)

load("../data/phy_3timepoints_restoration_score.rdata")

ps <- phy_3timepoints %>% subset_samples(Visit == "1 month" & !restoration_score %in% c(NA)) 
true_Rscore <- sample_df(ps) %>% dplyr::select(ABCNO, restoration_score)

sPLS_pred_1mon <- true_Rscore %>% full_join(Rscore_Preds_env_bac_1mon) %>% full_join(Rscore_Preds_env_1mon) %>% full_join(Rscore_Preds_bac_1mon)

library(ggplot2)
library(reshape2)

df_long <- melt(sPLS_pred_1mon, id.vars = c("ABCNO", "restoration_score"), variable.name = "sPLS_model", value.name = "Preds")
df_long$sPLS_model <- factor(df_long$sPLS_model, levels = c("RScore_Preds_bac_1mon","RScore_Preds_env_1mon","Rscore_Preds_env_bac_1mon"))

correlations <- lapply(sPLS_pred_1mon[,3:5], function(x) cor.test(x, sPLS_pred_1mon$restoration_score,method = "spearman",use = "complete.obs"))


cor_text <- data.frame(
  sPLS_model = names(correlations),
  cor_coeff = sapply(correlations, function(x) round(x$estimate,3)),
  p.value = sapply(correlations, function(x) x$p.value),
  x = c(0.42,0.42,0.42),
  y = c(0.24,0.205,0.215))

cor_text$p.value <- ifelse(cor_text$p.value <0.001," < 0.001", paste0(" = ",as.character(round(cor_text$p.value, 3))))
cor_text$label <- paste0("Spearman rho: ", round(cor_text$cor_coeff,2), ", P", cor_text$p.value)


p <- ggplot(df_long, aes(x = restoration_score, y = Preds, color = sPLS_model)) +
  geom_point(alpha = 0.9) + 
  geom_smooth(method = "lm", se = FALSE, aes(group = sPLS_model)) + 
  theme_minimal(base_size = 12) + 
  scale_color_manual(name = NULL,
                     values = c("Rscore_Preds_env_bac_1mon" = "#D55E00", "RScore_Preds_bac_1mon" = "#0072B2", "RScore_Preds_env_1mon" = "grey"),
                     labels = c(paste0("1 month gut microbiome",'\n', cor_text$label[which(cor_text$sPLS_model == "RScore_Preds_bac_1mon")]), 
                                paste0("Environmental factors",'\n', cor_text$label[which(cor_text$sPLS_model == "RScore_Preds_env_1mon")]),
                                paste0("1 month gut microbiome and environmental factors", '\n', cor_text$label[which(cor_text$sPLS_model == "Rscore_Preds_env_bac_1mon")])))+
  theme_minimal(base_size = 14) + 
  theme(legend.position = "bottom",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  labs(title = NULL, x = "Restoration score", y = "Predictions from sPLS models");p


plotname1 <- paste0(outpath, "/sPLS_1mon_corplot.png",sep = "")
ggsave(plotname1, p, width = 12, height = 9, limitsize = TRUE)

save(p, file = paste0(outpath, "/sPLS_1mon_corplot.RData"))


```







